Integrating brain and biomechanical models - a new paradigm for understanding neuro-muscular control

### 2017-11-08 papers/integration/omm_integ.tex
Adds changes based on Reviewer 1's comments on Q6, INTRODUCTION

% Solution described means we now need to do background of eye movements:
In an attempt to build a model combining brain, realistic
biomechanics \e{and} sensory feedback, we sought to extend our
previous work modelling the oculomotor system.  The existing
model \citep{cope_basal_2017} is already able to capture sensory input
and convert it into a neural signal, assumed to specify the target of
a \e{saccadic eye movement}; a fast movement of the eyes which directs
the fovea to a region of interest in the field of view.
%
The oculomotor system is an excellent candidate for modelling because
its movements can be specified with only three degrees of freedom,
making it one of the simplest neuro-muscular system in the body.
%
It is nevertheless behaviourally interesting, as saccadic eye
movements reveal information about decision making at a subconscious
level
\citep{deubel_saccade_1996,reppert_modulation_2015,marcos_determining_2016}.
The modelling of the oculomotor system is served by a large body of
behavioural data describing
saccades \citep{tipper_reaching_2001,walker_effect_1997,casteau_effect_2012},
many anatomical studies of the neural substrates involved
\citep{meredith_intrinsic_1998,isa_intrinsic_2002,isa_exploring_2009}
and electrophysiological data linking these together
\citep{hepp_spatio-temporal_1983,dorris_neuronal_1997,mcpeek_competition_2003,vokoun_circuit_2011}.
Furthermore, in the context of building \e{behaving} systems a
necessary part of any model for which the behaviour requires visual
attention and decision making is a realistic mechanism for gathering
visual information.  This is obvious from extrinsic considerations---a
subject must look at a scene to make decisions or navigate within
it. It also follows for \e{intrinsic} reasons. For
example, \cite{howard_hand_1997} showed that visual cues affect reach
trajectories and the same group later demonstrated that reaching
affects the saccadic system
\citep{tipper_reaching_2001} suggesting a close relationship between these
neural systems. Building the simplest, behaving oculomotor system will
therefore assist future computational neurobehavioural modelling
efforts.

Many neural populations are involved in the coding of saccadic eye
movements, only a very brief overview is given here; for a review,
see \cite{munoz_commentary:_2002}. One pathway takes information from
the retina directly into the superficial layers of the superior
colliculus in the brainstem
\citep{sterling_receptive_1971,linden_massive_1983,wu_involvement_1994}.
Activity within the superior colliculus then excites neurons in the
pons, medulla and rostral mid-brain \citep{sparks_brainstem_2002} and
then through channels to the motor neurons which innervate the
extraocular muscles
\citep{fuchs_firing_1970,sparks_brainstem_2002}. This direct pathway
is responsible for the low latency saccades called express saccades
\citep{schiller_effect_1987,edelman_activity_1996}.
%
Information from the retina is also processed by visual cortex which
feeds through to the frontal eye fields in which activity is related
to reflexive and voluntary
saccades \citep{schall_neural_1999}. Activity build-up in the frontal
eye fields is transferred to the intermediate layers of the superior
colliculus
\citep{stanton_frontal_1988-1} and is also processed by the basal ganglia,
which participates in the selection of the winning saccade end point
\citep{stanton_frontal_1988}.
%
Although both cortical and subcortical paths produce a saccade target
signal in the superior colliculus, it is also possible for animals to
make relatively normal saccades even after the colliculus has been
ablated
\citep{wurtz_activity_1972,aizawa_reversible_1998},
though express saccades are lost with collicular lesions
\citep{schiller_effect_1987}. This makes the superior colliculus
a perplexing structure, being both critically involved in saccade
target specification \citep{sparks_sensory_1987} and saccade dynamic control
\citep{waitzman_superior_1991,goossens_optimal_2012} and yet dispensible.
The `backup pathway' likely incorporates the oculomotor vermis and
fastigial oculomotor region of the cerebellum which are known to
participate in the specification, dynamics and adaptation of saccadic
eye movements \citep{kleine_saccade-related_2003,takagi_effects_1998}.

There is a long history of modelling the oculomotor system. For a
comprehensive review, see \cite{girard_brainstem_2005}. Models of
individual systems have been proposed for brainstem
\citep{robinson_oculomotor_1975,scudder_new_1988,gancarz_neural_1998},
cerebellum \citep{quaia_model_1999,dean_modelling_1995,dean_learning_1994}
and superior colliculus
\citep{arai_two-dimensional_1994,moren_mechanism_2013,marino_spatial_2012}.
More recently, combined models have also been developed incorporating
sensory input \citep{cope_basal_2017} and driving simple physical
plants representing the eye
\citep{tabareau_geometry_2007,nguyen_saccade_2014,thurat_biomimetic_2015}.
None of these models has yet fully closed the loop to produce a
behaving system operating freely within its environment. We argue that
developing integrated, closed-loop models of behaving systems offers
insights into the operation of neural systems that are not available
from input-~or output-assumption models.

@@

To the \ccg~model, we added a rate-coded brainstem model (a `saccadic
burst generator') and a biomechanical eye, implemented using the
biomechanical modelling framework OpenSim.  These will be described
below, but first we will give a description of the co-ordinate systems
and the modelling framework used.

@@

The D1MSN and D2MSN components are both leaky integrators, similar to
LINlinear. They differ in that they have no shunting inhibition. They
are used in to model medium spiny neuron (MSN) populations in the
striatum. As they model the fact that most MSN neurons fall into two
groups; those expressing D1 dopamine receptors and those expressing D2
receptors, they have a dopamine parameter that modulates the input
activation, so that their equations for $a_{in}$ are thus:
\begin{equation}
   a_{in}^{D1} = (0.2 + d)A + 0.01 R_N
\end{equation}
\begin{equation}
   a_{in}^{D2} = (1 - d)A + 0.01 R_N
\end{equation}
where $d$ is the dopamine parameter. Varying dopamine from 0 to 1
enhances the activation in the D1 model, whereas it decreases the
activation of the D2 model elements, in line with experimental
observations \citep{harsing_influence_1997,gonon_prolonged_1997}.
Note that the equation for $a_{in}^{D1}$ differs from that used in
the \ccg~model, for which the cortico-striatal weights are multiplied
by $(1+d)$ rather than $(0.2+d)$.

@@

We implemented a saccadic burst generator (SBG) based on the
connectivity outlined in \cite{gancarz_neural_1998}. The SBG network
for two of the model's six channels is shown in
Fig.~\ref{sbg}. Activity from the output layer of superior colliculus
(SC\_avg) is fed into each channel, which sums the activity it
receives and processes it in populations each of a single neural
element representing all the neurons in that population. Each channel
of the SBG functions to create the motor neuron activations that are
required to accelerate the eye in a particular direction, then hold
the eye in its new position against the returning force generated by
the elastic properties of the muscles.  The required motor neuron
activations are therefore a combination of features: a brief burst of
increased activity that accelerates the eye; followed by a period of
activity that is less than the burst firing rate but higher than the
tonic rate that exists when the eye is at the centre. This holds the
eye in its new position.

>>>>

% Solution described means we now need to do background of eye movements:
In an attempt to build a model combining brain, realistic
biomechanics \e{and} sensory feedback, we sought to extend our
previous work modelling the oculomotor system by adding a virtual,
biomechanical eye model able to make physically realistic
movements. The rotational state of the eye would then determine how
visual features in the virtual world were projected back into the
brain model.
%
The existing
model \citep{cope_basal_2017} is already able to capture sensory input
and convert it into a neural signal, assumed to specify the target of
a \e{saccadic eye movement}; a fast movement of the eyes which directs
the fovea to a region of interest in the field of view.
%
The oculomotor system is an excellent candidate for modelling because
its movements can be specified with only three degrees of freedom,
making it one of the simplest neuro-muscular systems in the body.
%
It is nevertheless behaviourally interesting, as saccadic eye
movements reveal information about decision making at a subconscious
level
\citep{deubel_saccade_1996,reppert_modulation_2015,marcos_determining_2016}.
The modelling of the oculomotor system is served by a large body of
behavioural data describing
saccades \citep{tipper_reaching_2001,walker_effect_1997,casteau_effect_2012},
many anatomical studies of the neural substrates involved
\citep{meredith_intrinsic_1998,isa_intrinsic_2002,isa_exploring_2009}
and electrophysiological data linking these together
\citep{hepp_spatio-temporal_1983,dorris_neuronal_1997,mcpeek_competition_2003,vokoun_circuit_2011}.
Furthermore, in the context of building \e{behaving} systems, a
necessary part of any model for which the behaviour requires visual
attention and decision making is a realistic mechanism for gathering
visual information.  This is obvious from extrinsic considerations---a
subject must look at a scene to make decisions or navigate within
it. It also follows for \e{intrinsic} reasons. For
example, \cite{howard_hand_1997} showed that visual cues affect reach
trajectories and the same group later demonstrated that reaching
affects the saccadic system
\citep{tipper_reaching_2001} suggesting a close relationship between these
neural systems. Building a behaving oculomotor system will
therefore assist future computational neurobehavioural modelling
efforts that involve reaching.

Many neural populations are involved in the coding of saccadic eye
movements, only a very brief overview is given here; for a review,
see \cite{munoz_commentary:_2002}. One pathway takes information from
the retina directly into the superficial layers of the superior
colliculus in the brainstem
\citep{sterling_receptive_1971,linden_massive_1983,wu_involvement_1994}.
Activity within the superior colliculus then excites neurons in the
pons, medulla and rostral mid-brain \citep{sparks_brainstem_2002}.
and finally the motor neurons which innervate the extraocular muscles
\citep{fuchs_firing_1970,sparks_brainstem_2002}. This direct pathway
is responsible for the low latency saccades called express saccades
\citep{schiller_effect_1987,edelman_activity_1996}.
%
Information from the retina is also processed by visual cortex which
feeds through to the frontal eye fields in which activity is related
to reflexive and voluntary
saccades \citep{schall_neural_1999}. Activity build-up in the frontal
eye fields is transferred to the intermediate layers of the superior
colliculus
\citep{stanton_frontal_1988-1} and is also processed by the basal ganglia,
which participates in the selection of the winning saccade end point
\citep{stanton_frontal_1988}.
%
Although both cortical and subcortical paths produce a saccade target
signal in the superior colliculus, it is also possible for animals to
make relatively normal saccades even after the colliculus has been
ablated
\citep{wurtz_activity_1972,aizawa_reversible_1998},
though express saccades are lost with collicular lesions
\citep{schiller_effect_1987}. This makes the superior colliculus
a perplexing structure, being both critically involved in saccade
target specification \citep{sparks_sensory_1987} and saccade dynamic control
\citep{waitzman_superior_1991,goossens_optimal_2012} and yet dispensible.
The `backup pathway' likely incorporates the oculomotor vermis and
fastigial oculomotor region of the cerebellum which are known to
participate in the specification, dynamics and adaptation of saccadic
eye movements \citep{kleine_saccade-related_2003,takagi_effects_1998}.

There is a long history of modelling the oculomotor system. For a
comprehensive review, see \cite{girard_brainstem_2005}. Models of
individual subsystems have been proposed for brainstem
\citep{robinson_oculomotor_1975,scudder_new_1988,gancarz_neural_1998},
cerebellum \citep{quaia_model_1999,dean_modelling_1995,dean_learning_1994}
and superior colliculus
\citep{arai_two-dimensional_1994,moren_mechanism_2013,marino_spatial_2012}.
More recently, combined models have also been developed incorporating
sensory input \citep{cope_basal_2017} and driving a second order
differential equation representing the eye
\citep{tabareau_geometry_2007,nguyen_saccade_2014,thurat_biomimetic_2015}.
None of these models has yet fully closed the loop to produce a
behaving system operating freely within its environment. We argue that
developing integrated, closed-loop models of behaving systems offers
insights into the operation of neural systems that are not available
from input-~or output-assumption models.

@@

To the \ccg~model, we added a rate-coded brainstem model (a `saccadic
burst generator') and a biomechanical eye, implemented using the
biomechanical modelling framework OpenSim.  These will be described
below, along with a review of the \ccg~model, but first we will give a
description of the co-ordinate systems and the modelling framework
used.

@@

The D1MSN and D2MSN components are both leaky integrators, similar to
LINlinear. They differ in that they have no shunting inhibition. They
are used to model medium spiny neuron (MSN) populations in the
striatum. As they model the fact that most MSN neurons fall into two
groups; those expressing D1 dopamine receptors and those expressing D2
receptors, they have a dopamine parameter that modulates the input
activation, so that their equations for $a_{in}$ are thus:
\begin{equation}
   a_{in}^{D1} = (0.2 + d)A + 0.01 R_N
\end{equation}
\begin{equation}
   a_{in}^{D2} = (1 - d)A + 0.01 R_N
\end{equation}
where $d$ is the dopamine parameter. Varying dopamine from 0 to 1
enhances the activation in the D1 model, whereas it decreases the
activation of the D2 model elements, in line with experimental
observations \citep{harsing_influence_1997,gonon_prolonged_1997}.
Note that the equation for $a_{in}^{D1}$ differs from that used in
the \ccg~model, for which the cortico-striatal weights are multiplied
by $(1+d)$ rather than $(0.2+d)$.

@@

We implemented a saccadic burst generator (SBG) based on the
connectivity outlined in \cite{gancarz_neural_1998}. The SBG network
for two of the model's six channels is shown in Fig.~\ref{sbg}. We use
the word `channel' to mean a set of populations of neurons which are
involved in actuating a single extraocular muscle). Activity from the
output layer of superior colliculus
%
(SC\_avg) is fed into each channel, which sums the activity it
receives and processes it in populations each of a single neural
element representing all the neurons in that population. Each channel
of the SBG functions to create the motor neuron activations that are
required to accelerate the eye in a particular direction, then hold
the eye in its new position against the returning force generated by
the elastic properties of the muscles.  The required motor neuron
activations are therefore a combination of features: a brief burst of
increased activity that accelerates the eye; followed by a period of
activity that is less than the burst firing rate but higher than the
tonic rate that exists when the eye is at the centre. This holds the
eye in its new position.
